---
title: "What Makes a Strong GDP"
subtitle: "Regression Analysis"
author: "Rtists"
date: "11/14/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries}
library(tidyverse)
library(readxl)
library(skimr)
```


```{r data, warning=FALSE}
economic_data <- read_excel("/cloud/project/02-data/economic_data.xlsx",
col_types = c("text", "text", "text",
"numeric", "numeric", "numeric",
"numeric", "numeric", "numeric",
"numeric", "numeric", "numeric",
"numeric", "numeric", "numeric"))
```


Your regression analysis results go here. At a minimum, the regression analysis should include the following: 

- Statement of the research question and modeling obejctive (prediction, inference, etc.)
- Description of the response variable
- Updated exploratory data analysis, incorporating any feedback from the proposal
- Explanation of the modeling process and why you chose those metohds, incorporating any feedback from the proposal
- Output of the final model
- Discussion of the assumptions for the final model
- Interpretations / interesting findings from the model coefficients
- Additional work of other models or analylsis not included in the final model.

*Use proper headings as needed.*

### Question and Objective

### Response Variable

### Exploratory Data Analysis

```{r inflation-categorize}
economic_data <- economic_data %>%
  mutate(cat_inflation = case_when(
Inflation >= 6 ~ "Dangerously High",
Inflation <6 & Inflation >3 ~ "High",
Inflation <= 3 & Inflation >=1 ~ "Healthy",
Inflation <1 & Inflation>0 ~ "Low",
Inflation <= 0 ~ "Deflation"
  ))
economic_data%>%
  group_by(cat_inflation) %>%
  count()
```


### Methods & Modelling

### Final Model

#### Create the model

#### Model Diagnostics
(influential points, leverage points)

### Discusson of Assumptions

### Interpretation and Other Interesting Findings

### Additional Information