---
title: "What Makes a Strong GDP?"
subtitle: "Regression Analysis"
author: "Rtists"
date: "11/14/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries}
library(tidyverse)
library(readxl)
library(skimr)
library(tidyr)
library(magrittr)
library(broom)
library(leaps)
library(knitr) 
library(rms)
```


```{r data, warning=FALSE}
economic_data <- read_excel("/cloud/project/02-data/economic_data.xlsx",
col_types = c("text", "text", "text",
"numeric", "numeric", "numeric",
"numeric", "numeric", "numeric",
"numeric", "numeric", "numeric",
"numeric", "numeric", "numeric"))
```


Your regression analysis results go here. At a minimum, the regression analysis should include the following: 

- Statement of the research question and modeling obejctive (prediction, inference, etc.)
- Description of the response variable
- Updated exploratory data analysis, incorporating any feedback from the proposal
- Explanation of the modeling process and why you chose those metohds, incorporating any feedback from the proposal
- Output of the final model
- Discussion of the assumptions for the final model
- Interpretations / interesting findings from the model coefficients
- Additional work of other models or analylsis not included in the final model.

*Use proper headings as needed.*

### Question and Objective

In this investigation, our goal is to try to understand what aspects of a country's economy are most associated with the GDP of the country. GDP is one of the most foundational measures of economic growth and the size of an economy, and as a result the majority of governments are extremely interested in understanding what contributes to having a large GDP. This study tries to help governments answer that question by investigating the research question: "What indicators of a country's economy have the greatest effect on the country's GDP?"

Since we already have data on all the countries and are thus don't need to try to "predict" the GDP of a country that may not be in our dataset, our modeling objective isn't prediction but rather inference, since we are trying to understand which predictor variable has the greatest effect on the response. Additionally, the modeling objective is inference as we are using our data set to build a model that reliably chooses the correct identifiers of GDP from a set of more variables than we may potentially need.

### Response Variable

In our investigation, the response value is GDP; that is, we are trying to determine the effect of various predictors on the GDP of countries. GDP represents the final value of all finished goods and services produced in a country in 2019, and in this case, in US dollars. However, according to the EDA conducted below, since the distribution of GDP's is incredibly right-skewed (countries such as the US and China have enormous GDP's matched by no other countries), we will be building the model using the log of GDP as our response.

### Exploratory Data Analysis

We will begin our exploratory data analysis by cleaning the data. We notice a total of 12 countries which have a "NA" value for one or more of the predictor variables that we will investigate. We will omit these entries from our analysis, but we expect this to add a slight bias to our results; the below 12 countries are all developing and thus our model may fit developed countries better than developing countries. However, the 12 countries below also have unusual governmental and economic structures as well as poor data quality associated with them; including them in the model may also significantly influence the coefficients associated with predictors. Thus, we have decided to leave them out of the model.

```{r NA-removal}
economic_data %>%
  filter(complete.cases(economic_data)==FALSE) %>%
  select(Country)
```
```{r omit-NA}
economic_data <- na.omit(economic_data)
```

### Response Variable

Now that the data has been cleaned, we can begin exploring our response variable.

```{r gdp-vis}
ggplot(data = economic_data, mapping = aes(x = GDP)) +
  geom_histogram(binwidth = 1000, fill = "forestgreen") +
  labs(x = "GDP (billions of USD)",
       y = "Frequency",
       title = "Distribution of GDPs")

economic_data %>%
select(GDP) %>%
skim()
```

The distribution of GDP shows significant right skew, which is reasonable and expected as the world has countries such as the US and China who, due to their population and industrial advantages, have significantly greater GDP's than the average country. We will now show a graph of GDP with some of these influential points removed.

```{r GDP without Outlier}
economic_data_temp <- economic_data %>% select(GDP) %>% filter(GDP < 3000)

ggplot(data = economic_data_temp, mapping = aes(x = GDP)) +
  geom_histogram(fill = "forestgreen") +
  labs(x = "GDP (billions of USD)",
       y = "Frequency",
       title = "Distribution of GDPs")

economic_data %>%
select(GDP) %>%
skim()
```

The distribution of GDP is unimodal and right-skewed. Since the media and interquartile range are less influenced by outliers, we report them as measures of center and spread. The median of the distribution is 88.9 billion and the interquartile range is 411.3. THe values are in billions of dollars.

This calls for a log-transform of the response variable, whose distribution is visualized below:

```{r log-gdp-vis}
ggplot(data = economic_data, mapping = aes(x = log(GDP))) +
  geom_histogram(binwidth = 1, fill = "forestgreen") +
  labs(x = "log(GDP)",
       y = "Frequency",
       title = "Distribution of log-GDPs")

economic_data <- economic_data %>%
  mutate(logGDP = log(GDP))

economic_data %>%
select(logGDP) %>%
skim()
```

This distribution of logGDP is normal and unimodal. We will likely use this as our response variable. Since there is minimal skewing and the distribution is generally symmetric, we report the mean and standard deviation as measures of center and spread. The mean logGDP is 4.61 and the standard deviation of its distribution is 2.08. 

### Predictor Variables

We will now look at each of the 12 predictor variables that we will use to predict GDP, starting with `TaxBurden`.

```{r Tax Burden}
ggplot(data = economic_data, mapping = aes(x = TaxBurden)) +
  geom_histogram(binwidth = 2, fill = "forestgreen") + 
  labs(x = "Tax Burden (% of Country's GDP)",
       y = "Frequency",
       title = "Distribution of Tax Burden")

economic_data %>%
select(TaxBurden) %>%
skim()
```

`TaxBurden` represents the amount of tax paid by the citizens of a country as a proportion of the GDP of that country. The distribution of tax burden is unimodal and only slightly right skewed. The mode is around 14-15%. In general, the tax burden across countries appears normally distributed. The mean tax burden is 22.19% and the standard deviation of the distribution is 10.17%. 

```{r Government Spending}
ggplot(data = economic_data, mapping = aes(x = GovSpending)) +
  geom_histogram(binwidth = 3, fill = "forestgreen") + 
  labs(x = "Government Spending (% of Country's GDP)",
       y = "Frequency",
       title = "Distribution of Government Spending")

economic_data %>%
select(GovSpending) %>%
skim()
```

`GovSpending` represents the amount spent by the government as a percentage of the GDP of the country. The distribution of government spending is generally symmmetric and unimodal. The mode of the distribution is around 25%. Since there is minimal skewing, we report the mean and standard deviation. The mean government spending is 33.87% of GDP and the distribution has a standard deviation of 15.52%.

```{r Population}
ggplot(data = economic_data, mapping = aes(x = Population)) +
  geom_histogram(binwidth = 40, fill = "forestgreen") + 
  labs(x = "Population (Million)",
       y = "Frequency",
       title = "Distribution of Population")

economic_data %>%
select(Population) %>%
skim()
```

`Population` represents the number of individuals living in a country. The distribution of population is unimodal and right-skewed. Because there are two influential points in "population", we will plot another graph of population without these two points below.

```{r Population without Outlier}
economic_data_temp <- economic_data %>% select(Population) %>% filter(Population < 500)

ggplot(data = economic_data_temp, mapping = aes(x = Population)) +
  geom_histogram(binwidth = 10, fill = "forestgreen") + 
  labs(x = "Population (Million)",
       y = "Frequency",
       title = "Distribution of Population")

economic_data %>%
select(Population) %>%
skim()
```

The distribution of population is unimodal and right-skewed. The mode of the distribution is around 1 million. Since the median and IQR are more robust to skewing, we report them instead as a measures of center and spread. The median is 9.15 and the IQR is 26.85. Additionally, when conducting our analysis, we may need to apply a log-transform to make the distribution of the variable more normal; based on a pairs plot, population has a logarithmic relationship with logGDP and thus we will apply a logarithmic transform to population below.

```{r logpop}
economic_data <- economic_data %>%
  mutate(logpop=log(Population))

ggplot(data=economic_data, aes(x=logpop)) +
  geom_histogram(fill = "forestgreen") +
  labs(title="Distribution of logpop", x="Log(Population)")
```

Based on the above plot, the logarithmic transformation has been successful in normalizing the population variable, which should allow a more linear relationship with GDP.

```{r Unemployment}
ggplot(data = economic_data, mapping = aes(x = Unemployment)) +
  geom_histogram(binwidth = 1, fill = "forestgreen") + 
  labs(x = "Unemployment (%)",
       y = "Frequency",
       title = "Distribution of Unemployment")

economic_data %>%
select(Unemployment) %>%
skim()
```

`Unemployment` represents the percantage of the workforce of a country that is currently not working. The distribution of unemployment is unimodal and right-skewed. The mode of the distribution is around 4-5%. Since the median and IQR are more robust to skewing, we report them instead as a measures of center and spread. The median is 5.7 and the IQR is 5.6.

```{r Inflation}
ggplot(data = economic_data, mapping = aes(x = Inflation)) +
  geom_histogram(fill = "forestgreen") + 
  labs(x = "Inflation (%)",
       y = "Frequency",
       title = "Distribution of Inflation")

economic_data_temp <- economic_data %>% select(Inflation) %>% filter(Inflation < 100)
ggplot(data = economic_data_temp, mapping = aes(x = Inflation)) +
  geom_histogram(binwidth = 1, fill = "forestgreen") + 
  labs(x = "Inflation (%)",
       y = "Frequency",
       title = "Distribution of Inflation without Outlier")

economic_data %>%
select(Inflation) %>%
skim()

economic_data <- economic_data %>%
  mutate(cat_inflation = case_when(
    Inflation >= 6 ~ "Dangerously High",
    Inflation <6 & Inflation >3 ~ "High",
    Inflation <= 3 & Inflation >=1 ~ "Healthy",
    Inflation <1 ~ "Low"))

economic_data%>%
  group_by(cat_inflation) %>%
  count()

economic_data <- economic_data %>%
  mutate(cat_inflation = as.factor(cat_inflation))

economic_data %>%
ggplot(mapping = aes(x = cat_inflation, y = logGDP)) +
  geom_boxplot(position="dodge",fill="forestgreen")
```

`Inflation` represents the change in prices of goods and services in a year in the country. The first distribution of inflation rates includes all outliers. Because of the size and resolution of the diagram, the distribution of most economies are not visible. The second diagram shows the distribution of inflation rates without the outlier with more than 1000% inflation. The distribution of inflation rate is generally unimodal and right skewed. The mode of the distribution is around 2%. Given significant skewing, we report the median and IQR as measures of center and spread. The median inflation rate is 2.7% and the IQR is 4%. Because inflation can vary widely between countries, we have recoded it into a categorical variable that better captures whether a country's inflation is healthy or unhealthy. Above is a plot to show the frequency of each level of inflation.


```{r Region}
regional <- economic_data %>%
  group_by(Region) %>%
  tally() %>%
  mutate(prop = n*100/nrow(economic_data))

ggplot(data = economic_data, mapping = aes(x = Region)) +
  geom_bar(fill = "forestgreen") + 
  labs(x = "Region",
       y = "Frequency",
       title = "Distribution of Region")

ggplot(mapping = aes(x = "", y = prop, fill = Region), data = regional) +
    geom_bar(stat = "identity", color = "white") +
    coord_polar("y", start = 0) +
    theme_void()
```

`Region` represents the geographical continent/area that the country is situated in. The bar graph and piechart of `Region` shows that there is a relatively equal representation of countries from different regions of the world. The Americas, Asia-Pacific, Sub-Saharan Africa, and Europe each represent around 25% of all the countries in the data. The smallest representation is from the Middle East and North Africa at 8.1%. We are not too concerned with the distribution because there are 195 countries in the world and our data has 173 countries. The difference in distribution across region is likely to be largely reflective of the actual geographical distribution of nation-states.

```{r Government Interference}
ggplot(mapping = aes(x = GovInterference), data = economic_data) +
  geom_bar(fill = "forestgreen") +
  labs(title = "Bar Graph of Government Inteference in Economy", x  = "Levels of Government Interference", y = "Frequency")
```

`GovInterference` represents the amount of interference that the government has in the economy as determined by the World Economic Index. The distribution of government interference shows that most countries either have extensive or moderate government inteference. The mode of the distribution is moderate government interference. Economies with limited and repressive intervention are significanly lesser, each having around 10-15 countries. We do not report the center or spread here because government interference is a categorical variable.

```{r mutate Government Interference}
economic_data <- economic_data %>%
  mutate(GovInterference = recode_factor(GovInterference, "Limited" =  "Moderate", "Repressive" = "Extensive"))

ggplot(mapping = aes(x = GovInterference), data = economic_data) +
  geom_bar(fill = "forestgreen") +
  labs(title = "Bar Graph of Government Inteference in Economy", x  = "Levels of Government Interference", y = "Frequency")
```

```{r Tariff Rate}
ggplot(mapping = aes(x = TariffRate), data = economic_data) + 
  geom_histogram(binwidth = 1, fill = "forestgreen") +
  labs(title = "Histogram of Tariff Rate", x  = "Tariff Rate", y = "Frequency")

economic_data %>%
  select(TariffRate) %>%
  skim()
```

`TariffRate` represents the average percentage tax on imports that the country has outstanding as of 2019. The distribution of tariff rate is generally right skewed and unimodal. There are several outlier economies with 50% tariff rate such as Central African Republic and North Korea. The mode of the distribution is around 2%. The median tariff rate is 4.3 and the interquartile range of the distribution is 6.7. Because of this right-skew, log-transforming this variable might be necessary; we will examine it further in a pairs plot later in the analysis.

```{r Income Tax Rate}
ggplot(mapping = aes(x = IncomeTaxRate), data = economic_data) +
  geom_histogram(binwidth = 5, fill = "forestgreen") +
  labs(title = "Histogram of Income Tax Rate", x  = "Income Tax Rate", y = "Frequency")

economic_data %>%
  select(IncomeTaxRate) %>%
  skim()
```

`IncomeTaxRate` represents the average tax rate applied to individuals on their incomes. The distribution of income tax rate is unimodal and generally symmetric. While its general shape resembles a normal distribution, there are several values of income tax rate which have particularly high frequency such as 10%, 25% and 34-35%. The mode of the distribution occurs at 35%. Since there is relatively minimal skewing, we report the mean and standard deviation as measures of center and spread. The mean income tax rate is 28.23 and the standard deviation of the distribution is 13.4. 

```{r Corporate Tax Rate}
ggplot(mapping = aes(x = CorporateTaxRate), data = economic_data) +
geom_histogram(binwidth = 4, fill = "forestgreen") +
labs(title = "Histogram of Corporate Tax Rate", x  = "Corporate Tax Rate", y = "Frequency")

economic_data %>%
select(CorporateTaxRate) %>%
skim()
```

`CorporateTaxRate` represents the average tax rate applied to corporations on their revenues. The distribution of corporate tax rate is unimodal and only slightly right skewed. The mode of the distribution is around 28-30%. Since there is minimal skewing, we report the mean and the standard deviation as measures of center and spread. The mean corporate tax rate is 23.89% and the standard deviation is 8.88%.

```{r Public Debt}
ggplot(data = economic_data, mapping = aes(x = PublicDebt)) +
  geom_histogram(binwidth = 8, fill = "forestgreen") + 
  labs(x = "Public Debt (% of GDP)",
       y = "Frequency",
       title = "Distribution of Public Debt")

economic_data %>%
select(PublicDebt) %>%
skim()
```

`PublicDebt` represents the debt of the country as a percentage of the country's GDP. The distribution of public debt is unimodal and right skewed. There are several outliers with public debt more than 175%. The mode of the distribution is around 30%. The median public debt is 49.4% and the interquartile range is 35%.

### Multivariate Exploration

Finally, we can visualize paired scatter plots of the relationship between GDP and each of our predictor variables. This is shown below. Many of the plots are visually difficult and don't offer good information as there are outliers that skew the visualization.

```{r scatterplot matrix GDP growth}
pairs(GDP ~ TariffRate + Population + Unemployment + Inflation + PublicDebt, data=economic_data, lower.panel = NULL)
pairs(GDP ~ GovSpending + IncomeTaxRate + CorporateTaxRate + TaxBurden, data=economic_data, lower.panel = NULL)
```

To better help us visualize the relationship between the predictor and response variables for most data points, we plotted a scatterplot matrix with most of the outliers removed.

```{r Scatterplot GDP without outliers}
temp <- economic_data %>%
  filter(GDP<3000 & Inflation < 100)

pairs(GDP ~ TariffRate + logpop + Population + Unemployment + Inflation + PublicDebt, data=temp, lower.panel = NULL)
pairs(GDP ~ GovSpending + IncomeTaxRate + CorporateTaxRate + TaxBurden, data=temp, lower.panel = NULL)
```

Based on the scatterplots above, we observe that most predictor variables do not have a clear linear relationship with the response variable GDP. Thus, we log transform GDP to see if there is a stronger linear relationship. As seen previously, logGDP has a normal distribution while GDP is extremely right skewed.

```{r Scatterplot logGDP}
pairs(logGDP ~ TariffRate + logpop + Population + Unemployment + Inflation + PublicDebt, data=temp, lower.panel = NULL)
pairs(logGDP ~ GovSpending + IncomeTaxRate + CorporateTaxRate + TaxBurden, data=temp, lower.panel = NULL)
```

Based on the scatterplots of the response variables against the quantitative predictor variables, we see that there is now a much stronger a linear relationship between the two. Particularly, we note that the logGDP against population scatterplot shows a curved distribution which resembles a logarithimic function. This is why we transformed population into the logpop variable, which has a linear relationship with GDP. Furthermore, the pairs plot validates all of the transformations we have applied in the univariate analysis, so they will remain in place for our regression.

Moreover, we also noticed some multicollinearity issues between predictor variables. In specific, the relationship between GovSpending and TaxBurden appears very strongly if not almost linear. Other variables suspect of multicollinearity problems include tariff rate and unemployment as well as tariff rate and unemployment.

```{r Boxplots for response vs categorical}
economic_data %>%
ggplot(mapping = aes(x = Region, y = logGDP)) +
  geom_boxplot(position="dodge") +
  labs(title = "Boxplots of log GDP versus Region")

economic_data %>%
ggplot(mapping = aes(x = GovInterference, y = logGDP)) +
  geom_boxplot(position="dodge") +
  labs(title = "Boxplots of log GDP versus Government Interference")
```

For the relationship between the response and categorical variables, the various boxplots suggest that there are generally normally distributed with minimal skewing. This means that a linear regression between log GDP and the categorical variables of region and government interference are appropriate.

We will now mean-center a few of the variables investigated thus far to ensure meaningful intercepts during the regression. 

```{r mean center variables}
economic_data <- economic_data %>%
  mutate(TaxBurdenCent = TaxBurden - mean(TaxBurden))
economic_data <- economic_data %>%
  mutate(GovSpendingCent = GovSpending - mean(GovSpending))
```


### Methods & Modelling

Our approach to model selection is as such: we will be running model selection using all three selection criteria - AIC, BIC, and adjusted R^2 - and thereafter compare the three models to decide on the final model.

We have three categorical variables here: cat_inflation, Region, and GovInterference. We begin by first setting "Healthy" as the baseline for cat_inflation and "Moderate" as the baseline for GovInterference. These are chosen as the baseline conditions of a healthy inflation rate and moderate government interference are often seen as the norm. Excessive government intervention and unhealthy inflation - whether is it too low, high, or dangerously high - are, in contrast, deviant conditions. 

```{r setting healthy as baseline, echo=TRUE}
economic_data <- economic_data %>%
  mutate(cat_inflation = relevel(cat_inflation, ref = "Healthy", ordered = FALSE))
```

```{r setting moderate as baseline, echo=TRUE}
economic_data <- economic_data %>%
  mutate(GovInterference = as.factor(GovInterference))

economic_data <- economic_data %>%
  mutate(GovInterference = relevel(GovInterference, ref = "Moderate", ordered = FALSE))
```


#### Selection using AIC

Below, we conduct model selection using AIC. We first use backward selection then forward selection and check if the final model selected is different.

```{r Modeling AIC backward, echo=TRUE}
aic_full_model <- lm(logGDP ~ TariffRate + logpop + Unemployment + cat_inflation + PublicDebt + GovSpendingCent + IncomeTaxRate + CorporateTaxRate + TaxBurdenCent + GDPGrowth + Region + GovInterference, data = economic_data)

aic_backward_model <- step(aic_full_model, direction = "backward")
```

```{r Modeling AIC forward, echo=TRUE}
aic_null_model <- lm(logGDP ~ 1, data = economic_data)

aic_forward_model <- step(aic_null_model, scope = formula(aic_full_model), direction = "forward")
```

AIC backward selection: logGDP ~ TariffRate + logpop + cat_inflation + GovSpendingCent + CorporateTaxRate + TaxBurdenCent + GDPGrowth + Region + GovInterference
    
AIC forward selection: logGDP ~ TariffRate + logpop + cat_inflation + GovSpendingCent + CorporateTaxRate + TaxBurdenCent + GDPGrowth + Region + GovInterference

As we have seen from the model selection process above, both forward and backward selection rendered the same final model. We will however not be choosing this model immediately. Instead, we will compare it with the other models generated using other selection criteria such as BIC and adj R^2.
    
```{r AIC model final}
tidy(aic_backward_model, conf.int="TRUE") %>%
  kable(format="markdown", conf.level = 0.95)
```
  
#### Selection using BIC

We now conduct model selection using BIC. Same as before, we do both forward and backward selection for comparison.

```{r BIC Model backward, echo=TRUE}
regfit_backward <- regsubsets(logGDP ~ TariffRate + logpop + Unemployment + cat_inflation + PublicDebt + GovSpendingCent + IncomeTaxRate + CorporateTaxRate + TaxBurdenCent + GDPGrowth + GovInterference + Region, data = economic_data, method="backward")

sel_summary <- summary(regfit_backward)
coef(regfit_backward, which.min(sel_summary$bic))
```

```{r BIC Model forward, echo=TRUE}
regfit_forward <- regsubsets(logGDP ~ TariffRate + logpop + Unemployment + cat_inflation + PublicDebt + GovSpendingCent + IncomeTaxRate + CorporateTaxRate + TaxBurdenCent + GDPGrowth + GovInterference + Region, data = economic_data, method="forward")

sel_summary <- summary(regfit_forward)
coef(regfit_forward, which.min(sel_summary$bic))
```

From the above, we see that both the forward and backward selection process rendered the same model.


#### Selection using adj R^2

We now conduct model selection using adj R^2. Same as before, we do both forward and backward selection for comparison.

```{r adjr2 Model backward, echo=TRUE}
sel_summary <- summary(regfit_backward)
coef(regfit_backward, which.max(sel_summary$adjr2))
```

```{r adjr2 Model forward, echo=TRUE}
sel_summary <- summary(regfit_forward)
coef(regfit_forward, which.max(sel_summary$adjr2))
```

From the above, we see that both the forward and backward selection process rendered the same model.


#### Model Comparison between all 3 Selection Criteria

Model selected using AIC: logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + GovSpendingCent + cat_inflationDangerouslyHigh + cat_inflationHigh + cat_inflationLow + RegionAsia-Pacific + RegionEurope + RegionMiddleEastandNorthAfrica + RegionSub-SaharanAfrica + GovInterferenceLimited + GovInterferenceExtensive + GovInterferenceRepressive
    
Model selected using BIC: logGDP ~  TariffRate + logpop + GDPGrowth + cat_inflationDangerouslyHigh + cat_inflationHigh + GovInterferenceExtensive + RegionSub-Saharan Africa             

Model selected using adj R^2: logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent +  cat_inflationDangerously High + cat_inflationHigh + GovInterferenceLimited + GovInterferenceModerate + RegionSub-Saharan Africa 

From the final equations generated using different selection criteria, we see that all three selection criteria included TariffRate + logpop + GDPGrowth + GovInterference in the final model. For that reason, we decide to include all four of these predictors into our final model.

cat_inflation is included in all three equations, though for BIC and adj R^2, the level of low inflation is excluded. This is not surprising considering that the p-value of the cat_inflationLow coefficient in the AIC model is 0.8030022, implying that this coefficient is probably very statistically insignificant. 

While there is one level that could be very statistically insignificant, we will choose to include cat_inflation in the final model for the following reasons. One, a categorical variable with 4 levels require us to use 3 dummy variables. Removing one because it is insignificant will affect the coefficients of the other levels because the missing level has to be forcibly added to either of the three category. We will also not change the categories of inflation rates because the present classification references economic literature which generally suggests four levels of inflation rates: low, healthy, high, and dangerously high (hyperinflation). Two, we also note that the AIC coefficient for cat_inflationLow to be extremely small. This means that the regression model only made very minimal changes to fit the data categorized as Low inflation, thus we are not too concerned that the inclusion of this level will significantly impact the accuracy of the regression. 

The same is observed for the categorical variable Region. The final BIC and adj R^2 models selected only RegionSub-SaharanAfrica. In this case, it is hard to conclude that the other levels of the categories are statistically insignificant as their p values in the final models are not extremely high/ close to 1. In fact, the coefficient of RegionMiddleEastandNorthAfrica has a very low p value at 0.0025057, while those of RegionEurope and RegionAsia-Pacific are above  From these circumstantial evidence, we are hesitant to recode the Region variable to have only two values: Sub-Saharan Africa and Not Subsaharan Africa. There seems to be a statistical difference between being in the Middle East and North Africa versus Europe or Asia Pacific.

GovSpendingCent is included only in the AIC equation while TaxBurdenCent is included in both the AIC and adjR2 equations. We conduct a nested F test to finally decide if these two predictor variables are significant to the model with predictor variables which we have decided to include thus far (TariffRate + logpop + GDPGrowth + cat_inflation + GovInterference + Region) based on the above reasoning.

Nested F test for GovSpendingCent:
```{r Nested F for GovSpendingCent, echo=TRUE}
reduced <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + cat_inflation + GovInterference + Region, data = economic_data)
full <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + cat_inflation + GovInterference + Region + GovSpendingCent, data = economic_data)
kable(anova(reduced, full), format="markdown", digits = 3)
```

The p value of the nested F test for GovSpendingCent is extremely high and thus we choose not include it in our model. Moreover, we also do not accord it high practical significance. The role of government spending is also highly contested in economic literature depending on which school of thought you subscribe to. That it is not a significant determinant of GDP reflect this ambuiguity in academic literature.


Nested F test for TaxBurdenCent:
```{r Nested F for TaxBurdenCent, echo=TRUE}
reduced <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + cat_inflation + GovInterference + Region, data = economic_data)
full <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + cat_inflation + GovInterference + Region + TaxBurdenCent, data = economic_data)
kable(anova(reduced, full), format="markdown", digits = 3)
```

The p value of the nested F test for TaxBurdenCent = 0.048 < 0.05. We thus choose to include TaxBurdenCent into our final model.


#### Testing Interaction Variables

To test for interaction between the categorical and predictor varaibles, we will conducted nested F tests. 

We do so by first testing for interaction between cat_inflation and the 4 quantitative variables.

Interaction between cat_inflation and TariffRate: 

```{r Nested F for TariffRate interaction with cat_inflation, echo=TRUE}
reduced <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region, data = economic_data)
full <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent+ cat_inflation + GovInterference + Region + cat_inflation*TariffRate, data = economic_data)
kable(anova(reduced, full), format="markdown", digits = 3)
```

Interaction between cat_inflation and logpop: 
```{r Nested F for logpop interaction with cat_inflation, echo=TRUE}
reduced <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region, data = economic_data)
full <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent+ cat_inflation + GovInterference + Region + cat_inflation*logpop, data = economic_data)
kable(anova(reduced, full), format="markdown", digits = 3)
```

Interaction between cat_inflation and GDPGrowth: 
```{r Nested F for GDPGrowth interaction with cat_inflation, echo=TRUE}
reduced <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region, data = economic_data)
full <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent+ cat_inflation + GovInterference + Region + cat_inflation*GDPGrowth, data = economic_data)
kable(anova(reduced, full), format="markdown", digits = 3)
```

Interaction between cat_inflation and TaxBurdenCent: 
```{r Nested F for TaxBurdenCent interaction with cat_inflation, echo=TRUE}
reduced <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region, data = economic_data)
full <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent+ cat_inflation + GovInterference + Region + cat_inflation*GDPGrowth, data = economic_data)
kable(anova(reduced, full), format="markdown", digits = 3)
```

As seen from the nested F tests, none of the p values are close to 0.05. This implies that none of the interaction between the 4 quantitive predictor variables and cat_inflation is statistically significant. 

We thus include no interaction variables for cat_inflation.


We now test for interaction between GovInterference and the 4 quantitative variables.

Interaction between GovInterference and TariffRate:
```{r Nested F for tariffrate interaction with GovInterference, echo=TRUE}
reduced <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region, data = economic_data)
full <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region + GovInterference*TariffRate, data = economic_data)
kable(anova(reduced, full), format="markdown", digits = 3)
```

Interaction between GovInterference and logpop:
```{r Nested F for logpop interaction with GovInterference, echo=TRUE}
reduced <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region, data = economic_data)
full <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region + GovInterference*logpop, data = economic_data)
kable(anova(reduced, full), format="markdown", digits = 3)
```

Interaction between GovInterference and GDPGrowth:
```{r Nested F for GDPgrowth interaction with GovInterference, echo=TRUE}
reduced <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region, data = economic_data)
full <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region + GovInterference*GDPGrowth, data = economic_data)
kable(anova(reduced, full), format="markdown", digits = 3)
```

Interaction between GovInterference and TaxBurdenCent:
```{r Nested F for TaxBurdenCent interaction with GovInterference, echo=TRUE}
reduced <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region, data = economic_data)
full <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region + GovInterference*TaxBurdenCent, data = economic_data)
kable(anova(reduced, full), format="markdown", digits = 3)
```

From the nested F tests above, we see that the p values for the interaction variables between GovInterference and TaxBurdenCent, GDPGrowth, logpop are all way above 0.05. Only the p value of the nested F test for the interaction between GovInterference and TariffRate is 0.067 which is close to 0.05. Since the p value is not far from 0.05 and we know that the dataset does not have a large number of observation, it is reasonable to assume that the effect of GovInterference on TariffRate must have been considerably large for the p value to be so close 0.05 even with a small dataset. 

Thus, we choose to include the interaction term GovInterference*TariffRate into the final model.


We now test for interaction between Region and the 4 quantitative variables.

Interaction between Region and TariffRate:
```{r Nested F for TariffRate interaction with region , echo=TRUE}
reduced <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region, data = economic_data)
full <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region + Region*TariffRate, data = economic_data)
kable(anova(reduced, full), format="markdown", digits = 3)
```

Interaction between Region and logpop:
```{r Nested F for logpop interaction with region , echo=TRUE}
reduced <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region, data = economic_data)
full <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region + Region*logpop, data = economic_data)
kable(anova(reduced, full), format="markdown", digits = 3)
```

Interaction between Region and GDPGrowth:
```{r Nested F for GDPGrowth interaction with region , echo=TRUE}
reduced <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region, data = economic_data)
full <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region + Region*GDPGrowth, data = economic_data)
kable(anova(reduced, full), format="markdown", digits = 3)
```

Interaction between Region and TaxBurdenCent:
```{r Nested F for taxburdenCent interaction with region , echo=TRUE}
reduced <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region, data = economic_data)
full <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region + Region*TaxBurdenCent, data = economic_data)
kable(anova(reduced, full), format="markdown", digits = 3)
```

From the nested F tests above, we see that the p values for the interaction variables between Region and TaxBurdenCent, GDPGrowth, logpop are all way above 0.05. Only the p value of the nested F test for the interaction between Region and TariffRate is 0.006 which is smaller than 0.05. This implies that there is significant interaction between Region and TariffRate.

Thus, we choose to include the interaction term Region*TariffRate into the final model.

### Final Model

We now regress with our final selection of predictor and interaction variables: TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region + GovInterference x TariffRate + Region x TariffRate.

```{r Final Model}
final_model <- lm(logGDP ~ TariffRate + logpop + GDPGrowth + TaxBurdenCent + cat_inflation + GovInterference + Region + GovInterference*TariffRate + Region*TariffRate, data=economic_data)
kable(tidy(final_model,conf.int=TRUE),format="html",digits=3)
```


#### Model Diagnostics

Next, we examine the four model assumptions: linearity, constant variance, normality, and independence. From previous scatterplots of response variable vs. each predictor variable, we didn't observe any non-linear relationships after we took a log transformation whenever necessary (for example, population). From the residuals vs. predicted values plot below, we can also conclude that the linearity assumption is satisfied. The following figure shows no obvious shape or pattern, and therefore, the constant variance assumption is also satisfied. 

```{r residual vs. predicted}
economic_data <- economic_data %>% 
  mutate(predicted = predict.lm(final_model), residuals = resid(final_model))

ggplot(data=economic_data,aes(x=predicted, y=residuals)) + 
  geom_point() + 
  geom_hline(yintercept=0,color="red") +
  labs(title="Residuals vs. Predicted Values")
```

```{r residuals vs. predictor variables}
ggplot(data=economic_data,aes(x=TariffRate, y=residuals)) + 
  geom_point() + 
  geom_hline(yintercept=0,color="red") +
  labs(x = "TariffRate", 
       y = "Residuals",
       title = "Scatterplot of Residuals vs. TariffRate")

ggplot(data=economic_data,aes(x=logpop, y=residuals)) + 
  geom_point() + 
  geom_hline(yintercept=0,color="red") +
  labs(x = "log of Population", 
       y = "Residuals",
       title = "Scatterplot of Residuals vs. log of Population")

ggplot(data=economic_data,aes(x=TaxBurdenCent, y=residuals)) + 
  geom_point() + 
  geom_hline(yintercept=0,color="red") +
  labs(x = "CorporateTaxRate", 
       y = "Residuals",
       title = "Scatterplot of Residuals vs. CorporateTaxRate")

ggplot(data=economic_data,aes(x=GovInterference, y=residuals)) + 
  geom_boxplot() + 
  geom_hline(yintercept=0,color="red") +
  labs(x = "GDPGrowth", 
       y = "Residuals",
       title = "Scatterplot of Residuals vs. GovInterference")

ggplot(data=economic_data,aes(x=cat_inflation,y=residuals)) + 
  geom_boxplot() + 
  geom_hline(yintercept=0,color="red") +
  labs(x = "Inflation", 
       y="Residuals")

ggplot(data=economic_data,aes(x=Region, y=residuals)) + 
  geom_boxplot() + 
  geom_hline(yintercept=0,color="red") +
  labs(x = "Region", 
       y = "Residuals",
       title = "Scatterplot of Residuals vs. Region")
```

The constant variance assumption is further satisfied because the residuals from each of the residuals vs. quantitative predictor variables plot is randomly scattered and doesn't appear any distinguishable shape. For those categorical predictor variables, the boxplot of residuals are appproximately centered around 0, sysmetric, and the interquantile range is roughly the same for each category. So, we can say that constant variance assumption is met.


```{r residual normality check}
ggplot(data = economic_data, aes(x = residuals)) +
  geom_histogram(binwidth = 0.2)+
  labs(x = "Residuals",
       y = "Frequency of Given Residual",
       title = "Histogram of Residuals")

ggplot(data=economic_data, aes(sample=residuals)) + 
  stat_qq() + 
  stat_qq_line() +
  labs(title="Normal QQ Plot of Residuals")
```

The normality assumption is satisfied since the distribution of residuals appears to be unimodal and approximately normal. Also, the normal QQ plot of points fit the theoretical line very well.

Lastly, we augmented the model and took a look into leverage, cook's distance, and standardized residual to check for outliers and assess their influence on our model.

```{r augment model}
economic_data_aug <- augment(final_model) %>%
  mutate(obs_num = row_number())

glimpse(head(economic_data_aug, n=5))
```

```{r leverage}
leverage_threshold <- 2*(8+1)/nrow(economic_data)

ggplot(data = economic_data_aug, aes(x = obs_num ,y = .hat)) + 
  geom_point(alpha = 0.7) + 
  geom_hline(yintercept = leverage_threshold,color = "red")+
  labs(x = "Observation Number",y = "Leverage",title = "Leverage") +
  geom_text(aes(label=ifelse(.hat > leverage_threshold, as.character(obs_num), "")), nudge_x = 4)

nrow(economic_data_aug %>% filter(.hat > leverage_threshold))

economic_data_aug %>% filter(.hat > leverage_threshold) %>%
  select(obs_num, .cooksd)
```

```{r cook distace}
ggplot(data = economic_data_aug, aes(x = obs_num, y = .cooksd)) + 
  geom_point(alpha = 0.7) + 
  geom_hline(yintercept=1,color = "red")+
  labs(x= "Observation Number",y = "Cook's Distance",title = "Cook's Distance") +
  geom_text(aes(label = ifelse(.hat>1,as.character(obs_num),"")))
```

There are 8 points with high leverage, but among all these points, none of them has a cook distance that is larger than 1. Therefore, we don't believe that they have significant influence on model coefficients, so it is safe to keep them in our model.

```{r standardized residuals}
ggplot(data = economic_data_aug, aes(x = .fitted,y = .std.resid)) +
  geom_point(alpha = 0.7) + 
  geom_hline(yintercept = 0,color = "red") +
  geom_hline(yintercept = -2,color = "red",linetype = "dotted") +
  geom_hline(yintercept = 2,color = "red",linetype = "dotted") +
  labs(x ="Predicted Value",y ="Standardized Residuals",title = "Standardized Residuals vs. Predicted") +
  geom_text(aes(label = ifelse(abs(.std.resid) >2,as.character(obs_num),"")), nudge_x = 0.3)

nrow(economic_data_aug %>% filter(abs(.std.resid) > 2))
```

There are 8 observations that are considered to have large standardized residuals with large magnitudes. If ploted the distribution of standardized residuals using a N(0,1) distribution, we can expect 8/173 = 4.62% of the observations have standardized residuals with magnitude > 2. We might need to remove these points from our final data in order to train a more accurate model. 

```{r VIF, echo=TRUE}
tidy(vif(final_model))
```

There doesn’t seem to be any multicollinearity in this model because the variance inflation factors are all below 10.

### Interpretation and Other Interesting Findings

Our final model can be represented by the following equation: 3.117 - .009(TariffRate) + 1.006(logpop) - .057(GDPGrowth) + .02(TaxBurdenCent) - .374(cat_inflationDangerously High) - .402(cat_inflationHigh) + .022(cat_inflationLow) - 1.045(GovInteferenceExtensive) + 1.117(RegionAsia-Pacific) + .537(RegionEurope) + .969(RegionMiddle East and North Africa) -.655(RegionSub-Saharan Africa) + .08(TariffRate:GovInteferenceExtensive) - .18(TariffRate:RegionAsia-Pacific) -.068(TariffRate:RegionEurope) + -.045(TariffRate:RegionMiddle East and North Africa) -.058(TariffRate:RegionSub-Saharan Africa)

*Each predictor can thus be interpreted in the following way:*
For countries who don't have "extensive" government interference and are in the Americas, as their TariffRate increases by 1%, the median GDP of the country is expected to be multiplied by a factor of `r exp(-.009)`.

For countries who are in the Asia Pacific, as their TariffRate increases by 1%, the median GDP of the country is expected to be multiplied by a factor of `r exp(-.18)`, holding all other variables constant.

For countries who are in Europe, as their TariffRate increases by 1%, the median GDP of the country is expected to be multiplied by a factor of `r exp(-.068)`, holding all other variables constant.

For countries who are in the Middle East or North Africa, as their TariffRate increases by 1%, the median GDP of the country is expected to be multiplied by a factor of `r exp(-.045)`, holding all other variables constant.

For countries who are in Sub-Saharan Africa, as their TariffRate increases by 1%, the median GDP of the country is expected to be multiplied by a factor of `r exp(-.058)`, holding all other variables constant.

For countries who do have extensive government interference, as the TariffRate increases by 1%, the median GDP of the country is expected to be multiplied by a factor of `r exp(.08)`, holding all other variables constant.

If the population is multiplied by a factor of C, the median GDP is expected to be multiplied by a factor of `exp(1.006logC)`.

As the Corporate Tax Rate increases by one percentage point, the expected median GDP of the country is multiplied by a factor of `r exp(-.036)`, holding all other variables constant. As the corporate tax rate increases, we can expect the GDP of the country to be lower.

As GDPGrowth increases by one percentage point, the expected median GDP of the country is multiplied by a factor of `r exp(-.057)`, holding all other variables constant. As GDP growth increases, we can expect the GDP of the country to be higher.

As the tax burden increases by one percentage point, the expected median GDP of the country is multiplied by a factor of `r exp(-.020)`, holding all other variables constant. As the tax burden increases, we can expect the GDP of the country to be lower.

If the country's inflation rate is listed as "Healthy", this provides no further input to the model as "Healthy" is the baseline for our model. However, if the country's inflation is listed as low, the expected median GDP of the country is multiplied by a factor of `r exp(.022)`, holding all other variables constant. If the country's inflation is listed as high, the expected median GDP of the country is multiplied by a factor of `r exp(-.402)`, holding all other variables constant. If the country's inflation is listed as dangerously high, the expected median GDP of the country is multiplied by a factor of `r exp(-.374)`, holding all other variables constant. 

If the country has extensive government interference, the expected median GDP of the country is multiplied by a factor of `r exp(-1.045)`, holding all other variables constant.

If the country is in the Americas, this provides no further input to the model as this is the baseline for our model. However, if the country is in the Asia Pacific region, the expected median GDP of the country is multiplied by a factor of `r exp(1.117)`, holding all other variables constant. If the country is in Europe, the expected median GDP of the country is multiplied by a factor of `r exp(.537)`, holding all other variables constant. If the country is in the Middle East or North Africa, the expected median GDP of the country is multiplied by a factor of `r exp(.969)`, holding all other variables constant. If the country is in Sub-Saharan Africa, the expected median GDP of the country is multiplied by a factor of `r exp(-.655)`, holding all other variables constant.

Beyond the specific interpretation of each of our coefficients, through the investigation we came across several interesting findings. The first and most notable is that none of our interaction effect terms involving inflation ended up in our final model as they did not have a significant effect on the response. We tried five different interaction effects with inflation: The effects between inflation and GDP growth, inflation and corporate tax rate, inflation and the log of the population, and inflation and the tariff rate. We hypothesized that the effect of these quantitative predictors may depend on the rate of inflation as this represents the stability of the currency which drives the country's economy, but this ended up not being the case. 

However, our hypothesis that the role of tariffs depended on region did end up being true as suggested by our generated p-values. We hypothesized that this was the case because depending on where the country is geographically positioned, tariffs as a means of protecting the country from foreign competition or disturbances could be very important or harmful. All of these interactions between tariff rate and region ended up in our final model.

