---
title: "PROJECT TITLE"
subtitle: "Project Proposal"
author: "NAME HERE"
date: "2019/10/26"
output: github_document
---

```{r libraries}
library(tidyverse)
library(readxl)
```

```{r load data}
setwd("/cloud/project")
economic_data <- read_excel("02-data/economic_data.xlsx", 
    col_types = c("text", "text", "text", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric"))

happiness_data <- read_excel("02-data/happiness_data.xlsx")

country_data <- full_join(economic_data, happiness_data, by = "Country")
country_data
```

## Section 1. Introduction

```{r glimpse}
glimpse(country_data)
```

We first take an overview of the data. We can see that there are 193 observations --- one observation corresponds to one country, and 16 variables, including country name, 14 predictor variables and one response variable that we want to predict, the happiness score.

## Section 2. Regression Analysis





## Section 3. Regression Analysis Plan

## Section 4. References

## The Data