---
title: "What makes a country economically great?"
subtitle: "Project Proposal"
author: "Rtists"
date: "2019/10/26"
output: github_document
---

```{r libraries}
library(tidyverse)
library(readxl)
library(skimr)
```


```{r data, warning=FALSE}
economic_data <- read_excel("/cloud/project/02-data/economic_data.xlsx",
col_types = c("text", "text", "text",
"numeric", "numeric", "numeric",
"numeric", "numeric", "numeric",
"numeric", "numeric", "numeric",
"numeric", "numeric", "numeric"))
```

## Section 1. Introduction

GDP is a measure of the total market value of all goods and services produced within a country over a period of time. Consequently, a high GDP implies that a country has significant economic influence. As emerging markets continue to industrialize, the question of how one achieves a larger GDP is becoming a critical one in policy selection; failure to design policies that encourage GDP growth can cause widespread suffering for a country's population. Consider the case of Venezuela, where GDP-friendly policies were shunned and massive unrest appeared as a nearly direct consequence. In order to avoid this fate, nations must analyze which economic sectors are most influential on GDP in order to create guiding policies. 

The use of multiple linear regression in the study of GDP and GDP growth is not novel. The approach has been widely employed by economists. For instance, Anghelache et. al has employed multiple linear regression model in analyzing the influence of final consumption and gross investment on Romania's GDP over time. Urrutia et. al, on the other hand, modeled Philippines's real GDP using multiple linear regression techniques. (Refer to citation of articles in the Citation section)

In this project, we will address the research question of what predictors are most useful in predicting GDP and to what extent GDP can be predicted successfully; we expect that factors such as population, inflation, GovInterference, and TaxBurden will most effectively predict GDP. The predictor variables used in our analysis are used in the calculation of the annual Economic Freedom Index. These predictors are not traditionally used in the multiple linear regression of GDP and we are interested to know the composite strength of these particular predictor variables. If we can get a better understanding of how these economic indicators correlate to GDP, it could provide policy makers and economists with a path towards sound governmental decisions. 

We hypothesize that out of the predictors in our dataset, `GovInterference` and both measures of `TaxRate` will have a very strong correlation with the GDP of the country, because they both reflect not only the government's economic policy but often times the government's approach to the macroecononmy. We predict that as government interference increases, GDP declines, and as tax rates increase, GDP declines as well.

## Section 2. Regression Analysis

```{r glimpse}
glimpse(economic_data)
```

From a glimpse of the data, can see that there are 193 observations --- each observation corresponds to one country has 15 variables providing information about it, including country name and 14 predictor variables that are economic and demographic indicators about the country. The data also contains the response variable that we want to predict, GDP. This response variable is a numeric and continuous variable that is in the billions of US dollars.

The data was collected by the World Economic Freedom Index, an organization that provides information and data about the economic health of countries and is used by policy makers to motivate their economic policy decisions.

Before looking at each of the predictor variables, we would like to ensure that the data is cleaned and error-free. We will remove observations that have missing information, such as countries like Iraq, Syria, North Korea, and others that have very little outside involvement. Additionally, we would like to remove the `GDPperCap` variable from the dataset, as it is simply the GDP divided by the population, both values we have as separate variables in the dataset, as to prevent potential future issues of multicolinearity.

```{r clean-data}
economic_data <- na.omit(economic_data)
economic_data$GDPperCap = NULL
```

Now that our data set is ready to explore, let's begin by looking at the distribution of GDP's of countries in our data set.

```{r gdp-vis}
ggplot(data = economic_data, mapping = aes(x = GDP)) +
  geom_histogram(binwidth = 1000) +
  labs(x = "GDP (billions of USD)",
       y = "Frequency",
       title = "Distribution of GDPs")
```

Clearly, there is incredible right skew in the data, which is reasonable as the world has countries such as the US and China with significantly greater GDP's than the average country. This calls for a log-transform of the response variable, whose distribution is visualized below:

```{r log-gdp-vis}
ggplot(data = economic_data, mapping = aes(x = log(GDP))) +
  geom_histogram(binwidth = 1) +
  labs(x = "log(GDP)",
       y = "Frequency",
       title = "Distribution of log-GDPs")
```

This distribution looks much more normal, and as a result, as we begin our analysis, we will likely use this as our response variable.

We will now look at each of the 12 predictor variables that we will use to predict GDP, starting with `TaxBurden`.

```{r Tax Burden}
ggplot(data = economic_data, mapping = aes(x = TaxBurden)) +
  geom_histogram(binwidth = 2) + 
  labs(x = "Tax Burden (% of Country's GDP)",
       y = "Frequency",
       title = "Distribution of Tax Burden")

economic_data %>%
select(TaxBurden) %>%
skim()
```

The distribution of tax burden is unimodal and only slightly right skewed. The mode is around 14-15%. In general, the tax burden across countries appear normally distributed. The mean tax burden is 22.19 and the standard deviation of the distribution is 10.17.

```{r Government Spending}
ggplot(data = economic_data, mapping = aes(x = GovSpending)) +
  geom_histogram(binwidth = 4) + 
  labs(x = "Government Spending (% of Country's GDP)",
       y = "Frequency",
       title = "Distribution of Government Spending")

economic_data %>%
select(GovSpending) %>%
skim()
```

The distribution of government spending is generally symmmetric and unimodal. There are several outliers which have significantly higher government spending: with government spending either equal to or more than 100% of the country's GDP. The mode of the distribution is around 25%. Since there is minimal skewing, we report the mean and standard deviation. The mean government spending is 33.87 and the distribution has a standard deviation of 15.52.

```{r Population}
ggplot(data = economic_data, mapping = aes(x = Population)) +
  geom_histogram(binwidth = 40) + 
  labs(x = "Population (Million)",
       y = "Frequency",
       title = "Distribution of Population")

economic_data %>%
select(Population) %>%
skim()
```

The distribution of population is unimodal and right-skewed.Because there are two extreme outliers in population, we will plot another graph of population without these two outliers below.

```{r Population without Outlier}
economic_data_temp <- economic_data %>% select(Population) %>% filter(Population < 500)

ggplot(data = economic_data_temp, mapping = aes(x = Population)) +
  geom_histogram(binwidth = 10) + 
  labs(x = "Population (Million)",
       y = "Frequency",
       title = "Distribution of Population")

economic_data %>%
select(Population) %>%
skim()
```

The distribution of population is unimodal and right-skewed. The mode of the distribution is around 1 million. Since the median and IQR are more robust to skewing, we report them instead as a measures of center and spread. The median is 9.15 and the IQR is 26.85.

```{r Unemployment}
ggplot(data = economic_data, mapping = aes(x = Unemployment)) +
  geom_histogram(binwidth = 1) + 
  labs(x = "Unemployment (%)",
       y = "Frequency",
       title = "Distribution of Unemployment")

economic_data %>%
select(Unemployment) %>%
skim()
```

The distribution of unemployment is unimodal and right-skewed. The mode of the distribution is around 4-5%. Since the median and IQR are more robust to skewing, we report them instead as a measures of center and spread. The median is 5.7 and the IQR is 5.6.

```{r Inflation}
ggplot(data = economic_data, mapping = aes(x = Inflation)) +
  geom_histogram() + 
  labs(x = "Inflation (%)",
       y = "Frequency",
       title = "Distribution of Inflation")

economic_data_temp <- economic_data %>% select(Inflation) %>% filter(Inflation < 100)
ggplot(data = economic_data_temp, mapping = aes(x = Inflation)) +
  geom_histogram(binwidth = 1) + 
  labs(x = "Inflation (%)",
       y = "Frequency",
       title = "Distribution of Inflation without Outlier")

economic_data %>%
select(Inflation) %>%
skim()
```
The first distribution of inflation rates includes all outliers. Because of the size and resolution of the diagram, the distribution of most economies are not visible. THe second diagram shows the distribution of inflation rates without the outlier with more than 1000% inflation. The distribution of inflation rate is generally unimodal and right skewed. The mode of the distribution is around 2%. Given significant skewing, we report the median and IQR as measures of center and spread. The median inflation rate is 2.7% and the IQR is 4%.

```{r Region}
aggregate(cbind(count = Region) ~ Region, data = economic_data, FUN = function(x){NROW(x)})

regional = data.frame("Region" = c("Americas","Asia-Pacific","Europe","Middle East and North Africa","Sub-Saharan Africa"), "share" = c(32, 43, 45, 18, 47), "prop"=c(17.4,23.2, 24.3, 9.7, 25.4))
mycols <- c("#0073C2FF", "#EFC000FF", "#868686FF", "#CD534CFF", "#99D492")

# Add label position
regional <- regional %>%
  arrange(desc(Region)) %>%
  mutate(lab.ypos = cumsum(prop) - 0.5*prop)
regional

ggplot(mapping = aes(x = "", y = share, fill = Region), data = regional) +
    geom_bar(stat = "identity", color = "white") +
    coord_polar("y", start = 0) +
  geom_text(aes(y = lab.ypos, label = prop), color = "white") +
    scale_fill_manual(values = mycols) +
    theme_void()
```

The piechart of region shows that there is a relatively equal representation of countries from different regions of the world. The Americas, Asia-Pacific, and Europe each represent around 25% of all the countries in the data. The smallest representation is from the Middle East and North Africa at 9.7%. We are not too concerned with the distribution because there are 195 countries in the world and our data has 185 countries. The difference in distribution across region is likely to be largely reflective of the actual geographical distribution of nation-states.

```{r Government Interference}
ggplot(mapping = aes(x = GovInterference), data = economic_data) +
geom_bar(fill = "cornflowerblue") +
labs(title = "Bar Graph of Government Inteference in Economy", x  = "Levels of Government Interference", y = "Frequency")
```

The distribution of government interference shows that most countries either have extensive or moderate government inteference. The mode of the distribution is moderate government interference. Economies with limited and repressive intervention are significanly lesser, each having around 10-15 countries. We do not report the center or spread here because government interference is a categorical variable.

```{r Tariff Rate}
ggplot(mapping = aes(x = TariffRate), data = economic_data) +
geom_histogram(fill = "cornflowerblue") +
labs(title = "Histogram of Tariff Rate", x  = "Tariff Rate", y = "Frequency")

economic_data %>%
select(TariffRate) %>%
skim()
```

The distribution of tariff rate is generally right skewed and unimodal. There are several outlier economies with 50% tariff rate such as Central African Republic and North Korea. The mode of the distribution is around 2%. The median tariff rate is 4.3 and the interquartile range of the distribution is 6.7. 

```{r Income Tax Rate}
ggplot(mapping = aes(x = IncomeTaxRate), data = economic_data) +
geom_histogram(fill = "cornflowerblue") +
labs(title = "Histogram of Income Tax Rate", x  = "Income Tax Rate", y = "Frequency")

economic_data %>%
select(IncomeTaxRate) %>%
skim()
```

The distribution of income tax rate is unimodal and generally symmetric. While its general shape resembles a normal distribution, there are several values of income tax rate which have particularly high frequency such as 10%, 25% and 34-35%. The mode of the distribution occurs at 35%. Since there is relatively minimal skewing, we report the mean and standard deviation as measures of center and spread. The mean income tax rate is 28.23 and the standard deviation of the distribution is 13.4. 

```{r Corporate Tax Rate}
ggplot(mapping = aes(x = CorporateTaxRate), data = economic_data) +
geom_histogram(fill = "cornflowerblue") +
labs(title = "Histogram of Corporate Tax Rate", x  = "Corporate Tax Rate", y = "Frequency")

economic_data %>%
select(CorporateTaxRate) %>%
skim()
```

The distribution of corporate tax rate is unimodal and only slightly right skewed. The mode of the distribution is around 28-30%. Since there is minimal skewing, we report the mean and the standard deviation as measures of center and spread. The mean corporate tax rate is 23.89% and the standard deviation is 8.88%.

```{r Public Debt}
ggplot(data = economic_data, mapping = aes(x = PublicDebt)) +
  geom_histogram(binwidth = 8) + 
  labs(x = "Public Debt (% of GDP)",
       y = "Frequency",
       title = "Distribution of Public Debt")

economic_data %>%
select(PublicDebt) %>%
skim()
```
The distribution of public debt is unimodal and right skewed. There are several outliers with public debt more than 175%. The mode of the distribution is around 30%. The median public debt is 49.4% and the interquartile range is 35%.

```{r scatterplot matrix GDP}
pairs((GDP) ~ TariffRate + Population + Unemployment + Inflation + PublicDebt, data=economic_data, lower.panel = NULL)

pairs(GDP ~ GovSpending + IncomeTaxRate + CorporateTaxRate + TaxBurden, data=economic_data, lower.panel = NULL)

temp <- economic_data %>%
  filter(GDP<3000)

pairs(GDP ~ TariffRate + Population + Unemployment + Inflation + PublicDebt, data=temp, lower.panel = NULL)

pairs(GDP ~ GovSpending + IncomeTaxRate + CorporateTaxRate + TaxBurden, data=temp, lower.panel = NULL)
```

```{r scatterplot matrix GDP growth}
pairs(GDPGrowth ~ TariffRate + Population + Unemployment + Inflation + PublicDebt, data=economic_data, lower.panel = NULL)

pairs(GDPGrowth ~ GovSpending + IncomeTaxRate + CorporateTaxRate + TaxBurden, data=economic_data, lower.panel = NULL)

temp2 <- economic_data %>%
  filter(GDPGrowth < 10, -10< GDPGrowth)

pairs(GDPGrowth ~ TariffRate + Population + Unemployment + Inflation + PublicDebt, data=temp2, lower.panel = NULL)

pairs(GDPGrowth ~ GovSpending + IncomeTaxRate + CorporateTaxRate + TaxBurden, data=temp2, lower.panel = NULL)

ggplot(data = temp, mapping = aes(x = TaxBurden, y = GDP)) +
  geom_point() + 
  labs(x = "Tax Burden (% of Country's GDP)",
       y = "Frequency",
       title = "Distribution of Tax Burden")
```

## Section 3. Regression Analysis Plan

## Section 4. References
https://www.heritage.org/index/ranking
https://ideas.repec.org/a/rsr/supplm/v61y2013i1p96-104.html
https://iopscience.iop.org/article/10.1088/1742-6596/820/1/012008
